---
title: "EDA"
output: github_document
---

```{r}
library(tidyverse) ##use ggplot 2
library(patchwork)
```

using dplyr verbs (group_by and summarize), along with numeric data summaries.


## Load NOAA weather dataset

Same dataset with Visualization part 1 and part 2, the only exception is the addition of month variable, created using lubridate::floor_date().

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

```

## group_by
Datasets are often comprised of groups defined by one or more (categorical) variable; group_by() makes these groupings explicit so that they can be included in subsequent operations. 

it is sometimes necessary to remove groups using ungroup().

use summarize to create one-number summaries within each group, or use mutate to define variables within groups.

```{r}
weather_df %>%
  group_by(name, month)
```

## Counting things

### Summarize
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(n_obs = n())
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(n_obs = n())
```

use summarize() to compute multiple summaries within each group. As an example, we count the number of observations in each month and the number of distinct values of date in each month.
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

### count()
To count things, you could use count() in place of group_by() and summarize() if you remember that this function exists. I’ll also make use of the name argument in count, which defaults to "n".

```{r}
weather_df %>%
  count(month, name = "n_obs")
```

count() is a useful tidyverse alternative to Base R’s table function. Both functions produce summaries of how often values appear, but table’s output is of class table and is hard to do any additional work with, while count produces a dataframe you can use or manipulate directly. For an example, run the code below and try to do something useful with the result…

```{r}
weather_df %>%
  pull(month) %>% 
  table

```

## (2x2 tables)
### group_by + summarize
for categorical variables:
look at the number of cold and not-cold days in Central Park and Waterhole.

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

### pivot_wider/janitor::tabyl

2x2 table using pivot_wider, or you could use janitor::tabyl
```{r}
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```

## General summarize
### mean(), median(), var(), sd(), mad(), IQR(), min(), and max()

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax),
    median_tmax = median(tmax))
```

### summarize multiple columns 
summarize multiple columns using the same summary, the across function is helpful.
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(across(tmin:prcp, mean))
```

### summerize + plot
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```

### summerize + pivot_wider

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)
```

## Group_by + mutate

Suppose you want to compare the daily max temperature to the annual average max temperature for each station separately, and to plot the result. 

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```

## Window functions
mean() takes n inputs and produces one single output.

Window functions, in contrast, take n inputs and return n outputs, and the outputs depend on all the inputs. 

### min_rank
find the max temperature ranking within month.
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax))
```

keep only the day with the lowest max temperature within each month:
```{r}
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(tmax) < 2)
```

keep the three days with the highest max temperature:
```{r}
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(desc(tmax)) < 4)
```
In both of these, we’ve skipped a mutate() statement that would create a ranking variable, and gone straight to filtering based on the result.

### Offsets

Offsets, especially lags, are used to compare an observation to it’s previous value. 

find the day-by-day change in max temperature within each station over the year:
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax))
```

### mean/sd
This kind of variable might be used to quantify the day-by-day variability in max temperature, or to identify the largest one-day increase:
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax)) %>%
  summarize(
    temp_change_sd = sd(temp_change, na.rm = TRUE),
    temp_change_max = max(temp_change, na.rm = TRUE))
```

## Limitations of Summerize 

summarize() can only be used with functions that return a single-number summary. This creates a ceiling, even if it is very high. Later we’ll see how to aggregate data in a more general way, and how to perform complex operations on the resulting sub-datasets.

## Revisiting examples

Learning Assessment: In the PULSE data, the primary outcome is BDI score; it’s observed over follow-up visits, and we might ask if the typical BDI score values are roughly similar at each. Try to write a code chunk that imports, cleans, and summarizes the PULSE data to examine the mean and median at each visit. Export the results of this in a reader-friendly format.
```{r}
pulse_data = 
  haven::read_sas("./public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

pulse_data %>% 
  group_by(visit) %>% 
  summarize(
    mean_bdi = mean(bdi, na.rm = TRUE),
    median_bdi = median(bdi, na.rm = TRUE)) %>% 
  knitr::kable(digits = 3) # exports it using knitr::kable.;
```

Learning Assessment: In the FAS data, there are several outcomes of interest; for now, focus on post-natal day on which a pup is able to pivot. Two predictors of interest are the dose level and the day of treatment. Produce a reader-friendly table that quantifies the possible associations between dose, day of treatment, and the ability to pivot.
```{r}
pup_data = 
  read_csv("./FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./FAS_litters.csv") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  group_by(dose, day_of_tx) %>% 
  drop_na(dose) %>% 
  summarize(mean_pivot = mean(pd_pivot, na.rm = TRUE)) %>% 
  pivot_wider(
    names_from = dose, 
    values_from = mean_pivot) %>% 
  knitr::kable(digits = 3)
```

