---
title: "EDA"
output: github_document
---

```{r}
library(tidyverse) ##use ggplot 2
library(patchwork)
```

using dplyr verbs (group_by and summarize), along with numeric data summaries.


## Load NOAA weather dataset

Same dataset with Visualization part 1 and part 2, the only exception is the addition of month variable, created using lubridate::floor_date().

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

```

## group_by
Datasets are often comprised of groups defined by one or more (categorical) variable; group_by() makes these groupings explicit so that they can be included in subsequent operations. 

it is sometimes necessary to remove groups using ungroup().

use summarize to create one-number summaries within each group, or use mutate to define variables within groups.

```{r}
weather_df %>%
  group_by(name, month)
```




