---
title: "Visualization Part1"
output: github_document
---

```{r}
library(tidyverse) ##use ggplot 2
library(ggridges)
```

## Load the dataset

NOAA weather data: use rnoaa::meteo_pull_monitors to download the dataset
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"), #specify the three weathersites
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"), 
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())


weather_df
```

##  Scatterplots

Creat first scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()

#OR 
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

### Save and edit a plot object
```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()

```
### color scatterplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) +
  geom_smooth(se = FALSE)

# OR
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE)

```

### add curve
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)


```

### seperate the scatterplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)

# OR
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .2) +
  geom_smooth(se = FALSE) + 
    facet_grid(. ~ name)
```
add curve
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

### density curve
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_density2d() +
  geom_point(alpha = .3) 
```

### some small notes

How many geoms have to exist?
You can have whatever geoms you want.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .2) +
  geom_smooth(se = FALSE) 
```


```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)
```

## Other plots
### Odds and ends

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)
```
There are lots of aesthetics, and these depend to some extent on the geom – color worked for both geom_point() and geom_smooth(), but shape only applies to points. The help page for each geom includes a list of understood aesthetics.

```{r}

ggplot(weather_df, aes(x = tmax, y = tmin)) + 
geom_bin2d()

ggplot(weather_df, aes(x = tmax, y = tmin)) +  
geom_density2d()
```

```{r}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
```
In the first attempt, we’re defining the color of the points by hand; in the second attempt, we’re implicitly creating a color variable that has the value blue everywhere; ggplot is then assigning colors according to this variable using the default color scheme.

## Histogram
```{r}

weather_df %>% 
  ggplot(aes(x = tmax)) + 
  geom_histogram()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram() +
  facet_grid(.~ name)
```

```{r}
weather_df %>% 
ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

## Density Plot

The adjust parameter in density plots is similar to the binwidth parameter in histograms, and it helps to try a few values. I set the transparency level to .4 to make sure all densities appear. You should also note the distinction between fill and color aesthetics here. You could facet by name as above but would have to ask if that makes comparisons easier or harder. Lastly, adding geom_rug() to a density plot can be a helpful way to show the raw data in addition to the density.

```{r}
weather_df %>% 
ggplot(aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

## Box Plot

```{r}
weather_df %>% 
ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

## Trendy plots
```{r}
weather_df %>% 
ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")

```

## Ridge plots
```{r}
weather_df %>% 
ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

