---
title: "Visualization Part2"
output: github_document
---

```{r}
library(tidyverse) ##use ggplot 2
library(patchwork)
```

## Load NOAA weather dataset

NOAA weather data: use rnoaa::meteo_pull_monitors to download the dataset
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())


weather_df
```

##  Review Scatterplots

Scatterplot of tmax against tmin

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5)
```

## Labels

Provide informative axis labels, plot titles, and captions, all of which can be controlled using labs().

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```

## Scale
坐标轴刻度线x轴y轴：scale_x_* and scale_y_* where * depends on the type of variable mapped to the x and y aesthetics (i.e. continuous vs discrete).

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15º C", "0", "15"))
```

### Complex Scales
scale_y_sqrt() can be added to a ggplot object to transform the Y scale
xlim() can be used to control the plot limits in the X axis.


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ºC", "0", "15"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```

arguments to scale_color_hue() control the color scale and the name in the plot legend.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
    scale_color_hue(name = "Location", h = c(100, 300))
```

use the viridis package to create your own color scheme 
The viridis::scale_fill_viridis() function is appropriate for the fill aesthetic used in histograms, density plots, and elsewhere

```{r}
ggp_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE  #discrete = TRUE because the color aesthetic is mapped to a discrete variable.if it is a continuous color gradient, omit this code;
  )

ggp_temp_plot

```

## Themes
Themes are used to modify non-data elements of a plot – they don’t change mappings or how data are render, but control things like background color and location of the the legend. Using themes can help with general plot appearance.

### legend position

```{r}
ggp_temp_plot + 
  theme(legend.position = "bottom")
```

remove ledend
```{r}
ggp_temp_plot + 
  theme(legend.position = "none")

```

### Different themes
While you can manage specific theme elements individually, I recommend using a built-in theme. By default this is theme_gray; 

here’s theme_bw():
Notes: The ordering of theme_bw() and theme() matters – theme() changes a particular element of the plot’s current “theme”. If you call theme to change the some element and then theme_bw(), the changes introduced by theme() are overwritten by theme_bw()
```{r}
ggp_temp_plot + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

theme_classic():
```{r}
ggp_temp_plot + 
  theme_classic() + 
  theme(legend.position = "bottom")
```

## Setting options
In addition to figure sizing, I include a few other figure preferences in global options declared at the outset of each .Rmd file 
this code chunk just gets copy-and-pasted to the beginning of every new file.
There are ways to set color preferences globally as well (for example, to use viridis color palettes.
```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data argument in geom_*

Depending on the setting, one way to do this is to create a “summary” dataframe and use that when adding a new geom to a ggplot based on the full data.
```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

